---
title: "How do people think the Covid-19 Pandemic will end?"
output: html_document
runtime: shiny
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(dplyr)
library(scales)
library(shiny)
library(summarytools)
setwd("D:\\Prerana Hiriyur\\Miscellaneous\\My projects\\Covid19 Perception\\Covid-19-End-Prediction-Survey")
data <- read.csv("Covid-19 Perceptions Survey (Responses) - Form responses.csv", header = TRUE)
data <- subset(data, select = -c(timestamp))
```

### An analysis of responses from a survey on Covid-19 predictions

The Covid-19 pandemic has caused an unimaginable amount economic, social and medical damage all over the world and naturally most of us have been anxious to know how and when it will come to an end. As discussing multiple opinions and hypotheses over FaceTime and Zoom calls soon got cliche and boring, I decided to conduct a short survey asking people to predict the most likely ending to the pandemic. In this document, I present an analysis of the responses to this survey.

# Results {.tabset}

```{r data, include=FALSE}
pred.end.levels <- c("The virus will mutate and become mild or harmless in the next few months", 
                     "An effective vaccine will be widely available within 4-6 months", 
                     "Without further lockdown, this virus will spread rapidly in the next 2-4 months to infect most of the population causing herd immunity", 
                     "With intermittent lockdown and releases, this virus will linger on it its current form for the next 1-2 years", 
                     "I do not see an end to this pandemic in the foreseeable future. This virus is here to stay.")
pred.end.labels <- c("Mutation", "Effective Vaccine", "Herd Immunity", "Prolonged Recovery", "No End")
data$predicted.end.label <- factor(data$predicted.end, 
                                   levels=pred.end.levels, 
                                   labels=pred.end.labels)
profession.levels <- c("Healthcare (Doctor/Nurse/Paramedic)", 
                       "Healthcare related (Pharmaceuticals/Biotechnology/Biomedical Engineering, etc)", 
                       "Non-healthcare related profession", 
                       "Currently not working")
profession.labels <- c("Healthcare", 
                       "Healthcare related", 
                       "Non-healthcare", 
                       "Not working")
data$profession.label <- factor(data$profession, levels = profession.levels, labels=profession.labels)
covid.stats.levels <- c("I am keeping track of real time Covid-19 data", 
                        "I check every day", 
                        "I check once every few days", 
                        "I check once a while (weekly/bi-weekly)",
                        "I don't pay attention to numbers")
covid.stats.labels <- c("Real Time", 
                        "Daily", 
                        "Once in a few days", 
                        "Weekly or Bi-weekly", 
                        "Anti-Stats")

data$covid19.stats.label <- factor(data$covid19.stats, covid.stats.levels, covid.stats.labels)
data$primary.source.of.news <- factor(data$primary.source.of.news, 
                                      levels = c("Television", 
                                                 "News apps/ Internet", 
                                                 "Social Media", 
                                                 "Newspaper", 
                                                 "Other"))
data$preferred.sm.platform <- factor(data$preferred.sm.platform, 
                                      levels = c("Facebook", 
                                                 "Instagram", 
                                                 "Twitter", 
                                                 "Whatsapp", 
                                                 "Other"))
edu.levels <- c("High school graduate, diploma or the equivalent",
                "Bachelor’s degree", 
                "Master’s degree", 
                "Professional degree", 
                "Doctorate degree")
edu.labels <- c("High School", 
                "Bachelor's Degree", 
                "Master's degree", 
                "Professional degree", 
                "Doctorate degree")
data$highest.level.of.edu.label <- factor(data$highest.level.of.edu, edu.levels, edu.labels)

```


## Descriptive Stats {.tabset}

The dataset contains 1066 responses to the survey which can be found [here](https://docs.google.com/forms/d/e/1FAIpQLSeUBxDmHKE8G2ui8Nz0Z2KdPmT2QOa-LhXT-fu2nyRjzB3HcA/viewform?usp=sf_link). I used my parents' Whatsapp and Facebook accounts as a channel to gather a large number of responses from middle-aged adults and my own to gather responses from younger adults. A more detailed description of the dataset can be found in the following tabs. 

### Dataset Summary

```{r, echo=FALSE, results='asis', message=FALSE}
Dataset <- data
dfSummary(Dataset, plain.ascii = FALSE, style = "grid", headings = FALSE, tmp.img.dir = "\\tmp")

```

### Graphs

```{r, fig.height = 5, fig.width = 8, fig.align = "center", echo=FALSE, results='asis', message=FALSE}

selectInput("graph", "Choose a variable:", 
            c("Age"="age", 
              "Gender"="gender",
              "Country" = "country", 
              "Level of Education" = "leveledu", 
              "Profession" = "profession", 
              "Comorbid Conditions" = "comorbid", 
              "Infected by Covid-19" = "infectedcovid",
              "Primary Source of News" = "primarynews", 
              "Preferred Social Media Platform" = "preferredsm",
              "Social Media Activity" = "smactivity",
              "Closest Known Covid19 Case" = "closestcovid",
              "Frequency of tracking Covid-19 statistics" = "covidstats", 
              "Preferred Healthcare System" = "preferredhealth", 
              "Belief in Vaccination" = "beliefinvaccine",
              "Predicted End Scenario" = "predictedend"), 
            selected = NULL, multiple = FALSE, selectize = TRUE)


renderPlot({
  gender <- ggplot(data, aes(gender)) + 
              geom_bar(width = 0.6, alpha=0.8, aes(fill=factor(gender))) +
                  theme(legend.position = "none", plot.title = element_text(hjust = 0.5)) +
                  geom_text(stat='count', aes(label=..count..), vjust=-0.3) +
                  xlab("") + 
                  ylab("Frequency") + 
                  ggtitle("Gender") 
  age <- ggplot(data, aes(age)) + 
              geom_histogram(aes(y=..density..), fill="#00BFC4", alpha=0.7) +
                  scale_x_continuous(breaks = scales::pretty_breaks(n = 10)) +
                  theme(plot.title = element_text(hjust = 0.5)) +
                  labs(title = "Age", x = "Age", y = "Density") +
                  scale_x_continuous(breaks = c(seq(0,90,10)))
  profession <- ggplot(data, aes(profession.label)) + 
                    geom_bar(width = 0.6, alpha=0.8, aes(fill=factor(profession.label))) +
                    theme(legend.position = "none", plot.title = element_text(hjust = 0.5)) +
                    geom_text(stat='count', aes(label=..count..), vjust=-0.3) +
                    xlab("") + 
                    ylab("Frequency") + 
                    ggtitle("Profession")
  covidstats <- ggplot(data, aes(covid19.stats.label)) + 
                    geom_bar(width = 0.6, alpha=0.8, aes(fill=factor(covid19.stats.label))) +
                    theme(legend.position = "none", plot.title = element_text(hjust = 0.5)) +
                    geom_text(stat='count', aes(label=..count..), vjust=-0.3) +
                    xlab("") + 
                    ylab("Frequency") + 
                    ggtitle('How often do you check Covid-19 statistics?')
  primarynews <- ggplot(data, aes(primary.source.of.news)) + 
                    geom_bar(width = 0.6, alpha=0.8, aes(fill=factor(primary.source.of.news))) +
                    theme(legend.position = "none", plot.title = element_text(hjust = 0.5)) +
                    geom_text(stat='count', aes(label=..count..), vjust=-0.3) +
                    xlab("") + 
                    ylab("Frequency") + 
                    ggtitle("Primary Source of News") 
  predictedend <- ggplot(data, aes(predicted.end.label)) + 
                    geom_bar(width = 0.6, alpha=0.8, aes(fill=factor(predicted.end.label))) +
                    theme(legend.position = "none", plot.title = element_text(hjust = 0.5)) +
                    geom_text(stat='count', aes(label=..count..), vjust=-0.3) +
                    xlab("") + 
                    ylab("Frequency") + 
                    ggtitle("How do you think Covid-19 will end?") 
  preferredhealth <- ggplot(data, aes(preferred.healthcare.sys)) + 
                        geom_bar(width = 0.6, alpha=0.8, aes(fill=factor(preferred.healthcare.sys))) +
                        theme(legend.position = "none", plot.title = element_text(hjust = 0.5)) +
                        geom_text(stat='count', aes(label=..count..), vjust=-0.3) +
                        xlab("") + 
                        ylab("Frequency") + 
                        ggtitle("Preferred Healthcare System") 
  closestcovid <- ggplot(data, aes(closest.covid.case.label)) + 
                    geom_bar(width = 0.6, alpha=0.8, aes(fill=factor(closest.covid.case.label))) +
                    theme(legend.position = "none", plot.title = element_text(hjust = 0.5)) +
                    geom_text(stat='count', aes(label=..count..), vjust=-0.3) +
                    xlab("") + 
                    ylab("Frequency") + 
                    ggtitle("Closest Known Covid-19 Case")
  smactivity <- ggplot(data, aes(sm.activity)) + 
                  geom_bar(width = 0.6, alpha=0.8, aes(fill=factor(sm.activity))) +
                  theme(legend.position = "none", plot.title = element_text(hjust = 0.5)) +
                  geom_text(stat='count', aes(label=..count..), vjust=-0.3) +
                  xlab("") + 
                  ylab("Frequency") + 
                  ggtitle("Social Media Activity") 
  preferredsm <- ggplot(data, aes(preferred.sm.platform)) + 
                    geom_bar(width = 0.6, alpha=0.8, aes(fill=factor(preferred.sm.platform))) +
                    theme(legend.position = "none", plot.title = element_text(hjust = 0.5)) +
                    geom_text(stat='count', aes(label=..count..), vjust=-0.3) +
                    xlab("") + 
                    ylab("Frequency") + 
                    ggtitle("Preferred Social Media Platform")
  infectedcovid <- ggplot(data, aes(infected)) + 
                      geom_bar(width = 0.6, alpha=0.8, aes(fill=factor(infected))) +
                      theme(legend.position = "none", plot.title = element_text(hjust = 0.5)) +
                      geom_text(stat='count', aes(label=..count..), vjust=-0.3) +
                      xlab("") + 
                      ylab("Frequency") + 
                      ggtitle("Have you been infected by Covid-19?") 
  comorbid <- ggplot(data, aes(comorbid)) + 
                geom_bar(width = 0.6, alpha=0.8, aes(fill=factor(comorbid))) +
                theme(legend.position = "none", plot.title = element_text(hjust = 0.5)) +
                geom_text(stat='count', aes(label=..count..), vjust=-0.3) +
                xlab("") + 
                ylab("Frequency") + 
                ggtitle("Do you or anyone in your family have comorbid conditions?")
  leveledu <- ggplot(data, aes(highest.level.of.edu.label)) + 
                geom_bar(width = 0.6, alpha=0.8, aes(fill=factor(highest.level.of.edu.label))) +
                theme(legend.position = "none", plot.title = element_text(hjust = 0.5)) +
                geom_text(stat='count', aes(label=..count..), vjust=-0.3) +
                xlab("") + 
                ylab("Frequency") + 
                ggtitle("Level of Education") 
  country <- ggplot(data, aes(country)) + 
              geom_bar(width = 0.6, alpha=0.8, aes(fill=factor(country))) +
              theme(legend.position = "none", plot.title = element_text(hjust = 0.5)) +
              geom_text(stat='count', aes(label=..count..), vjust=-0.3) +
              xlab("") + 
              ylab("Frequency") + 
              ggtitle("Country")
  beliefinvaccine <- ggplot(data, aes(belief.in.vaccination)) + 
                        geom_bar(width = 0.6, alpha=0.8, fill = "#F8766D") +
                        theme(legend.position = "none", plot.title = element_text(hjust = 0.5)) +
                        geom_text(stat='count', aes(label=..count..), vjust=-0.3) +
                        xlab("") + 
                        ylab("Frequency") + 
                        ggtitle("Belief in Vaccination") +
                        scale_x_continuous(breaks=c(seq(0,5))) + 
                        scale_y_continuous(breaks = c(seq(0,600,100)))
    x = list("age" = age, 
              "gender" = gender,
              "country" = country, 
              "leveledu" = leveledu, 
              "profession" = profession, 
              "comorbid" = comorbid, 
              "infectedcovid" = infectedcovid,
              "primarynews" = primarynews, 
              "preferredsm" = preferredsm,
              "smactivity" = smactivity,
              "closestcovid" = closestcovid,
              "covidstats" = covidstats, 
              "preferredhealth" = preferredhealth, 
              "beliefinvaccine" = beliefinvaccine,
              "predictedend" = predictedend)
  
  x[input$graph]
})

```



## Hypotheses {.tabset}

My inspiration for this survey came from a conversation I was having with my dad about the state of the Covid-19 pandemic. My father is a doctor and I wanted to know his perspective on the discourse about vaccines, lockdowns, etc. Specifically, we were discussing the various vaccine trials and "breakthroughs" reported on the news and in social media. While it seemed to me that a vaccine may be discovered soon, my dad seemed to believe that even a safe and widely distributed vaccine is very unlikely to be verry effective. He drew a comparison between the nature of the corona virus and the influenza virus. The unimpressive efficacy of the flu vaccine, when compared to vaccines for small pox and measles, was his basis for believing that the most likely end to the pandemic would be the achievement of herd immunity. 

This made me wonder if profession (mainly healthcare vs non-healthcare industries) along with other factors would affect an individual's prediction about the end to this pandemic. So I created this short survey to try and test hypotheses about the impact of profession, social media usage, exposure to Covid-19 statistics, gender, etc. on predictions about how the pandemic is most likely to end. 

The tabs below contain a few of the hypotheses that I found to be most interesting. Some of the other possible hypotheses are listed under "Others". 

### Profession

Naturally, the first question I wanted to ask was **whether non-healthcare professionals more optimistic about the discovery of an effective vaccine in the near future than the healthcare professionals?**

The graph below shows the proportion of healthcare professionals who chose the Vaccine Scenario as opposed to the other 3 profession categories.

```{r fig2, fig.height = 5, fig.width = 8, fig.align = "center", echo=FALSE}
proptable.prof.end = as.data.frame(table(data$profession.lab, data$predicted.end.lab))

proptable.prof.end = proptable.prof.end %>%
  rename(
    Profession = Var1,
    Scenario = Var2
  ) %>%
  group_by(Profession) %>%
  mutate(Proportion = Freq / sum(Freq)) %>%
  mutate(Percentage = round(Proportion*100, 2))

ggplot(proptable.prof.end, aes(fill=Scenario, y=Proportion, x=Profession)) +
  geom_bar(position = "fill", stat="identity") +
  geom_text(aes(label = paste0(round(Percentage,2),"%")), 
            position = position_stack(vjust = 0.5), size = 3) +
  scale_y_continuous(labels = scales::percent) +
  xlab("")

```

From the figure above, there is a clear difference (21.89% vs 34.29%) in the proportion of individuals who predicted the discovery of an effective vaccine in "Healthcare" and "Non-healthcare" categories. Those who are not currently working seem to be all the more optimistic (which I will explore in the later tab called Gender), along with those in "Healthcare related" professions although the number of respondents in this category is too small to make conclusions. 

To be sure that the differences in proportions in "Healthcare" and "Non-healthcare" are significant and not due to randomness in the data, I did a **chi-squared test**

```{r ttest1, echo=FALSE}
freq.vaccine.healthcare <- as.integer(count(data[which(data$profession.label=="Healthcare" &
                                              data$predicted.end.label=="Effective Vaccine"),]))
freq.vaccine.nonhealthcare <- as.integer(count(data[which(data$profession.label=="Non-healthcare" &
                                          data$predicted.end.label=="Effective Vaccine"),]))
total.healthcare <- as.integer(count(data[which(data$profession.label=="Healthcare"),]))
total.nonhealthcare <- as.integer(count(data[which(data$profession.label=="Non-healthcare"),]))
test = prop.test(c(freq.vaccine.healthcare, freq.vaccine.nonhealthcare), c(total.healthcare,total.nonhealthcare))

print(test)
```

The p-value (0.001~) is extremely small, therefore we can be confident that there is in fact a significant difference in the proportions. 

While this conclusion seems to make sense, it could be **confounded** by other characteristics. One common confounder in most analyses is age. In the following plots I explore the relationship between age, prediction and profession. 

```{r ageconfound, echo=FALSE, fig.align="center"}
ggplot(data, aes(x=profession.label, y=age)) + 
  geom_boxplot(aes(fill=factor(profession.label))) + 
  theme(legend.position = "none", plot.title = element_text(hjust = 0.5)) +
  labs(x = "", y = "Age")
```

The **box plot** above shows that the average age among healthcare professionals is higher than non-healthcare professionals. If age is also correlated with the prediction, it could turn out to be a confounder.


The following **jitter plot** helps us understand the relationship between age and prediction better. If age were correlated with the prediction, we would see patches of dots at different locations in the plot. 

```{r confound2, echo=FALSE, fig.align="center"}
ggplot(data, aes(y=age, x=predicted.end.label, color=predicted.end.label)) +
  geom_jitter() + 
  theme(legend.position = "none", plot.title = element_text(hjust = 0.5)) +
  labs(x = "", y = "Age", title = "Jitter Plot of Age by Chosen Scenario")
```

Here we see that the large patches are mostly evenly distributed except for a concentration of dots at the bottom of the "Effective Vaccine" category, suggesting that among the small number of young participants, there is a higher expectation of a Vaccine. While this could possibly affect the above conclusion, the number in this cluster is quite small and therefore the effect is likely to be negligible. 




### Social Media 

Another hypothesis I considered was **if social media activity had an effect on the perception of how the pandemic will end**. The question in the survey elicited information on frequency of social media usage on a scale of 1 to 5, 1 being "Almost never go online" to 5, "Go online almost every hour". In the section on Descriptive Statistics, you can see the overall response to this question. As expected, very few respondents chose 1 and 2, the most frequent option is 3 followed by 4 and 5. 

The stacked columns below display the proportions of ending scenarios chosen by each category. 

```{r fig.align="center", fig.height = 5, fig.width = 8, echo=FALSE}
proptable.sm.activity = as.data.frame(table(data$sm.activity, data$predicted.end.lab))

proptable.sm.activity = proptable.sm.activity %>%
  rename(
    SM.Activity = Var1,
    Scenario = Var2
  ) %>%
  group_by(SM.Activity) %>%
  mutate(Proportion = Freq / sum(Freq)) %>%
  mutate(Percentage = round(Proportion*100, 2))

ggplot(proptable.sm.activity, aes(fill=Scenario, y=Proportion, x=SM.Activity)) +
  geom_bar(position = "fill", stat="identity") +
  geom_text(aes(label = paste0(round(Percentage,2),"%")),
            position = position_stack(vjust = 0.5), size = 2) +
  scale_y_continuous(labels = scales::percent) +
  labs(x = "Social Media Activity")

```
From the figure, it seems as though individuals who use social media very frequently (5) are more likely to have a negative perception about the ending of the pandemic (Option 5: "I see no end to the pandemic in the near future"). Although the proportion seems large among groups 1 and 2, these groups are too small to make any conclusive statements about them. 

The following chi-squared test on proportions determines if the difference in proportion of people choosing scenario 5 in the frequent social media usage category vs others is significant or not. 

```{r, echo=FALSE}
freq.vaccine.sm5 <- as.integer(count(data[which(data$sm.activity==5 &
                                                         data$predicted.end.label=="No End"),]))
freq.vaccine.smnot5 <- as.integer(count(data[which(data$sm.activity!=5 &
                                                            data$predicted.end.label=="No End"),]))
total.sm5 <- as.integer(count(data[which(data$sm.activity==5),]))
total.smnot5 <- as.integer(count(data[which(data$sm.activity!=5),]))
test = prop.test(c(freq.vaccine.sm5, freq.vaccine.smnot5), c(total.sm5,total.smnot5))
print(test)
```

As the p-value (0.027) is below 5% we can reject the null in favor of the alternative (difference is significant).

Once again, we must still check for confounders. Again one possible confounder is age, the following box plot compares the distribution of age among the 5 categories. 

```{r fig.align="center", fig.height = 5, fig.width = 8, echo=FALSE}
ggplot(data, aes(y=age, x=sm.activity, fill=sm.activity)) +
  geom_boxplot(aes(group=sm.activity)) + 
  theme(legend.position = "none", plot.title = element_text(hjust = 0.5)) +
  labs(x = "Social Media Activity", y = "Age", title = "Jitter Plot of Age by Social Media Activity") 
```
The distribution among groups 3,4 and 5 that we are concerned about (large enough number of responses) does not seem very different. 

Of course, age does affect the choice of platform. The following jitter plot shows the very obvious difference in platform preferences among age groups. If the choice of platform affects perception, this could be a possible confounder too. 


```{r fig.align="center", fig.height = 5, fig.width = 8, echo=FALSE}
ggplot(data, aes(y=age, x=preferred.sm.platform, color=preferred.sm.platform)) +
  geom_jitter() + 
  theme(legend.position = "none", plot.title = element_text(hjust = 0.5)) +
  labs(x = "", y = "Age", title = "Jitter Plot of Age by Social Media Platform")

```
From the jitter plot above, it is evident that the most popular platform is Whatsapp. In fact, it dominates all other choices (making up 64.2% of the data). By analyzing the responses of just Whatsapp users, we can double check our earlier inference. 

```{r fig.align="center", fig.height = 5, fig.width = 8, echo=FALSE}
data_whatsapp = subset(data, preferred.sm.platform=="Whatsapp")
ggplot(data_whatsapp, aes(sm.activity)) + 
  geom_bar(width = 0.6, alpha=0.8, aes(fill=factor(sm.activity))) +
  theme(legend.position = "none", plot.title = element_text(hjust = 0.5)) +
  geom_text(stat='count', aes(label=..count..), vjust=-0.3) +
  xlab("Social Media Activity") + 
  ylab("Frequency") + 
  ggtitle("Social Media Activity Among Whatsapp Users") 
```
The distribution of social media activity levels among whatsapp users seems similar to the whole dataset. 


```{r fig.align="center", fig.height = 5, fig.width = 8, echo=FALSE}
proptable.sm.activity.w = as.data.frame(table(data_whatsapp$sm.activity, data_whatsapp$predicted.end.label))

proptable.sm.activity.w = proptable.sm.activity.w %>%
  rename(
    SM.Activity = Var1,
    Scenario = Var2
  ) %>%
  group_by(SM.Activity) %>%
  mutate(Proportion = Freq / sum(Freq)) %>%
  mutate(Percentage = round(Proportion*100, 2))

ggplot(proptable.sm.activity.w, aes(fill=Scenario, y=Proportion, x=SM.Activity)) +
  geom_bar(position = "fill", stat="identity") +
  geom_text(aes(label = paste0(round(Percentage,2),"%")),
            position = position_stack(vjust = 0.5), size = 2) +
  scale_y_continuous(labels = scales::percent) +
  labs(x = "Social Media Activity")

```
Once again from the plot we can see a similar pattern, in fact a more stark difference in the proportion of individuals choosing Scenario 5 among highly frequent users of Social Media. 

```{r, echo=FALSE}
freq.vaccine.sm5 <- as.integer(count(data_whatsapp[which(data_whatsapp$sm.activity==5 &
                                                         data_whatsapp$predicted.end.label=="No End"),]))
freq.vaccine.smnot5 <- as.integer(count(data_whatsapp[which(data_whatsapp$sm.activity!=5 &
                                                            data_whatsapp$predicted.end.label=="No End"),]))
total.sm5 <- as.integer(count(data_whatsapp[which(data_whatsapp$sm.activity==5),]))
total.smnot5 <- as.integer(count(data_whatsapp[which(data_whatsapp$sm.activity!=5),]))
test = prop.test(c(freq.vaccine.sm5, freq.vaccine.smnot5), c(total.sm5, total.smnot5))
print(test)
``` 

The chi-squared test above with p-value (0.024) confirms this. 

### Gender

I also wanted to check **if gender had an effect on response to the final question** of the survey. From the descriptive statistics we can see that the number of male and female respondents are roughly equal. The following column chart shows the proportion of end scenarios chosen between male and female respondents. 

```{r fig.align="center", fig.height = 5, fig.width = 8, echo=FALSE}
proptable.gender = as.data.frame(table(data$gender, data$predicted.end.lab))

proptable.gender = proptable.gender %>%
  rename(
    Gender = Var1,
    Scenario = Var2
  ) %>%
  group_by(Gender) %>%
  filter(Gender != 'Prefer not to say') %>%
  mutate(Proportion = Freq / sum(Freq)) %>%
  mutate(Percentage = round(Proportion*100, 2))

ggplot(proptable.gender, aes(fill=Scenario, y=Proportion, x=Gender)) +
  geom_bar(position = "fill", stat="identity") +
  geom_text(aes(label = paste0(round(Percentage,2),"%")),
            position = position_stack(vjust = 0.5), size = 2) +
  scale_y_continuous(labels = scales::percent) +
  labs(x = "")
```
The choices seem to be similar between the 2 groups with female respondents being only slightly more optimistic about an effective vaccine than men. 

The following column chart shows the proportion of males and females in the profession categories that we found had a significant impact on the prediction. 

```{r fig.align="center", fig.height = 5, fig.width = 8, echo=FALSE}
proptable.gender.prof = as.data.frame(table(data$gender, data$profession.label))

proptable.gender.prof = proptable.gender.prof %>%
  rename(
    Gender = Var1,
    Profession = Var2
  ) %>%
  filter(Gender != 'Prefer not to say') %>%
  group_by(Profession) %>%
  mutate(Proportion = Freq / sum(Freq)) %>%
  mutate(Percentage = round(Proportion*100, 2))

ggplot(proptable.gender.prof, aes(fill=Gender, y=Proportion, x=Profession)) +
  geom_bar(position = "fill", stat="identity") +
  geom_text(aes(label = paste0(round(Percentage,2),"%")),
            position = position_stack(vjust = 0.5), size = 2) +
  scale_y_continuous(labels = scales::percent) +
  labs(x = "")

```
The distribution of male and female are similar in all groups except the "Not Working" category. In my earlier analysis, I only drew a conclusion regarding the differences in optimism about a vaccine in the "Healthcare" and "Non-Healthcare" categories. The chart below shows a much larger optimism in the "Not Working" category which may contain some gender effects along with the effect of profession. 

```{r fig.align="center", fig.height = 5, fig.width = 8, echo=FALSE}
ggplot(proptable.prof.end, aes(fill=Scenario, y=Proportion, x=Profession)) +
  geom_bar(position = "fill", stat="identity") +
  geom_text(aes(label = paste0(round(Percentage,2),"%")), 
            position = position_stack(vjust = 0.5), size = 2) +
  scale_y_continuous(labels = scales::percent) 
```

### Others 

#### **Frequency of Tracking Covid-19 Statistics**

**Another possible effect on prediction could be the frequency of tracking case numbers**. While tracking numbers regularly could mean better awareness, at the time of this survey, India was seeing a continuous, fast-paced, daily increase in numbers. Tracking this too often could also amount to fear. I wanted to check the differences in prediction based on this behaviour.  


The following stacked column chart shows the proportions of different predictions chosen by individuals in different tracking behaviour categories. From the Descriptive Statistics tab, we can see that the number of respondents in the "Real Time" tracking category is very low and therefore it would be best to focus only on the other 4 groups. 

```{r fig.align="center", fig.height = 5, fig.width = 8, echo=FALSE}
proptable.stats.end = as.data.frame(table(data$covid19.stats.label, data$predicted.end.lab))

proptable.stats.end = proptable.stats.end %>%
  rename(
    Rate = Var1,
    Scenario = Var2
  ) %>%
  group_by(Rate) %>%
  mutate(Proportion = Freq / sum(Freq)) %>%
  mutate(Percentage = round(Proportion*100, 2))

ggplot(proptable.stats.end, aes(fill=Scenario, y=Proportion, x=Rate)) +
  geom_bar(position = "fill", stat="identity") +
  geom_text(aes(label = paste0(round(Percentage,2),"%")), 
            position = position_stack(vjust = 0.5), size = 2) +
  scale_y_continuous(labels = scales::percent) +
  labs(x = "") 
```

Although there seem to be differences in proportions across these groups, the pattern of such difference is not clear. FOr example, there seems to be an increasing trend in the proportion choosing the "Mutation" scenario across the groups. Those who are less exposed to statistics seem to be more optimistic in this respect. However there is also an increasing trend in the "No End" Scenario which seems to contradict that conclusion. So to make better conclusions would require more information, either in the form of more questions about where people are getting their statistics from, whether they are aware of inaccuracies in reporting of numbers in India, etc. 



</br>

#### **Primary Source of News**

Since information about the spread of Covid-19 is widely available on various media formats, I wanted to check **if the different media formats leave different impressions on their users**. The following column chart represents this interaction.

```{r fig.align="center", fig.height = 5, fig.width = 8, echo=FALSE}
proptable.news.end = as.data.frame(table(data$primary.source.of.news, data$predicted.end.lab))

proptable.news.end = proptable.news.end %>%
  rename(
    PrimarySource = Var1,
    Scenario = Var2
  ) %>%
  group_by(PrimarySource) %>%
  mutate(Proportion = Freq / sum(Freq)) %>%
  mutate(Percentage = round(Proportion*100, 2))

ggplot(proptable.news.end, aes(fill=Scenario, y=Proportion, x=PrimarySource)) +
  geom_bar(position = "fill", stat="identity") +
  geom_text(aes(label = paste0(round(Percentage,2),"%")), 
            position = position_stack(vjust = 0.5), size = 2) +
  scale_y_continuous(labels = scales::percent) +
  labs(x = "") 

```

From the figure above, it seems like Television viewers are more optimistic than those who get their news from apps and social media. A larger proportion of TV viewers chose Scenarios 1 and 2 (Mutation and Effective Vaccine). It is possible that this finding is confounded by age.

It would be interesting to know if regional, national and international news sources have different impacts too and in hindsight would add this as another question in the survey. 


## Summary

I conducted this survey to understand how people think the on-going pandemic is most likely to end and understand if certain factors strongly influence their perception. Unsurprisingly, I found that fewer medical professionals believe in the possibility of an effective vaccine compared to other profession categories. I also found that social media activity affects perception as well, with highly frequent users being more likely to see no end to current crisis. I also found that in this dataset, women were slightly more optimistic about the discovery of a vaccine. 

I am aware of some of the main limitations to this analysis. One obvious one being a distribution of age that is not representative of the general Indian population. The average age in this dataset is 45, while India has a much younger age distribution. This is mostly because of the method I used to circulate this survey. It is also a very small dataset mostly eliciting responses from residents in Bangalore (a large city in India). Of course, residents of smaller towns, less drastically affected by the pandemic may have a different perception. 

This was my first attempt to conduct a survey and analyse the results and I learned a lot. I gained a better understanding of how to frame questions, how to present results and how to acknowledge the limitations of anonymous survey data. I hope you enjoyed going through this project as much as I enjoyed working on it! 



# {-}
